---
title: "WUP_04"
author: "Luz Vargas"
date: "Fall 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
data = read.csv('GSScsv.csv')
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey. The question to answer is to resolve this hypothesis.\
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling EDUC\
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of EDUC

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We try to predict WORDSUM based on degree.\
WORDSUM is numerical and Degree is categorical therefore we will use the Numerical Categorical method.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
boxplot(WORDSUM~DEGREE,data=data)
```


Based on the evidence in the Boxplot, Graduate and Less than High School have the highest average difference in WORDSUM scores.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=data)
```


The numerical summary shows that Graduate's have the highest median at 11, Bachelor's and Junior College have the second highest median at 10, the third highest is High School at 9, and Lower then High School has the lowest median at 8.  

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfies the assumptions and is appropriate to answer the research question!-->


```{r}
model1 <- aov(WORDSUM~DEGREE,data=data)
summary.aov(model1)
```
Since the p-value is 0.000236 we reject the null hypothesis because the p-value is less than 0.05.  


```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 17, 2, 2))
plot(TukeyHSD(model1),las=2)
```

Since High School and Lt High School do not pass 0, that means there is a difference in average score on the WORDSUM test between at least two levels of EDUC.  


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->



